Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-08 paper.tex
conclusion

The neural recognition model accelerates inference,
but given enough iterations of our algorithm,
a lesioned version without the recognition model eventually comes close to catching up (green curves in Figure~\ref{regressionLearningCurve}; No NN in Table~\ref{regressionComparison}).
We compare our model on held-out tasks against 4 baselines (Table~\ref{regressionComparison}).
\begin{figure}\centering
  \includegraphics[width = 7cm]{figures/regression.eps}
  \caption{Learning curves for symbolic regression. Red: \system. Green: no neural network. Notice how the model without the neural network eventually solves almost all the tasks, but does so using programs with lower likelihood (see Table~\ref{regressionComparison}).}\label{regressionLearningCurve}
\end{figure}

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string routines provided by Python --
and automatically recover many of the higher-level building blocks that have made these
other system successful.
Table~\ref{textPrimitives} lists the primitive functions provided to \system.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide array/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).

We automatically generated  600 string editing along with input/output examples (Figure~\ref{exampleTextProblem}), split  75/25 train/test.

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly can find those solutions.
For symbolic regression, we also care about the quality
of the solution, as measured by the likelihood model $\probability[x|p]$.
Concretely, this cashes out as measuring how many
real numbers were used in the programs: did the model correctly explain a linear function
using two numbers, or did it introduce extraneous parameters?
Table~\ref{baselineComparisons}
compares our model against these baselines.

\begin{table}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours&Ours--NN  & RF/DC & PCFG & Enum

@@

  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&99\% &82\% \\
  Solve time&  0.9s&2.2s
  \\\bottomrule
  \end{tabular}
\caption{\% solved: fraction solved with 5 sec timeout. Solve time: averaged over solved tasks. MDL: $-\expect\left[\probability[x|p] \right]$. For domains other than symbolic regression MDL is 0 nats. For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}
\pagebreak
 \section{Related Work}
 Our work is far from the first to learn to learn programs,
 an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:

@@

 \noindent \emph{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and go beyond it along two dimensions: (1) we propose more sophisticated algorithms for
 inducing reusable structure, based on a formalism called \emph{Fragment Grammars}~\cite{tim};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} and Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}.

@@

 \section{Conclusion}
 We contributed a model, \system, that learns to program by
 bootstrapping a domain-specific language.
 The  DSL learns to expose new domain-specific primitives
  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the desired programs.  We believe that this
 integration of top-down symbolic representation and bottom-up neural
 networks -- both of them learned -- is a useful general approach for
 getting program induction off the ground.

>>>>

%% The neural recognition model accelerates inference,
%% but given enough iterations of our algorithm,
%% a lesioned version without the recognition model eventually comes close to catching up (green curves in Figure~\ref{regressionLearningCurve}; No NN in Table~\ref{regressionComparison}).
%% We compare our model on held-out tasks against 4 baselines (Table~\ref{regressionComparison}).

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string routines provided by Python --
and automatically recover many of the higher-level building blocks that have made these
other system successful.
Table~\ref{textPrimitives} lists the primitive functions provided to \system.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).

We automatically generated  600 string editing tasks along with input/output examples (Figure~\ref{exampleTextProblem}). %, split  75/25 train/test.

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
For symbolic regression, we also care about the quality
of the solution, as measured by the likelihood model $\probability[x|p]$.
Concretely, this cashes out as measuring how many
real numbers were used in the programs: e.g. did the model correctly explain a linear function
using two numbers, or did it introduce extraneous parameters?
Table~\ref{baselineComparisons}
compares our model against these baselines.
We see that our full model consistently
beats the baselines,
and that the recognition model 
consistently increases the number of solved held-out tasks.
Lesioning the recognition model also slows down the convergence of the algorithm,
taking more wake/sleep cycles to reach a given number of tasks solved (Figure~\ref{learningCurves}).
This data supports our view of the recognition model as a way of accelerating the search over programs.
\begin{table}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours&Ours--NN  & RF/DC & PCFG & Enum

@@

  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&99\% &82\% \\
  Solve time&  0.9s&2.2s
  \\\bottomrule
  \end{tabular}
\caption{\% solved: fraction solved with 5 sec timeout. Solve time: averaged over solved tasks. MDL: $-\expect\left[\probability[x|p] \right]$. For domains other than symbolic regression MDL is 0 nats. For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}

\begin{figure}\centering
  \includegraphics[width = 7cm]{figures/regression.eps}
  \caption{Learning curves for \system both with and without the recognition model (No NN in the above plots). Red: \system. Green: no neural network. Notice how the model without the neural network eventually solves almost all the tasks, but does so using programs with lower likelihood (see Table~\ref{regressionComparison}).}\label{regressionLearningCurve}  
\end{figure}

\pagebreak
 \section{Related Work}
 Our work is far from the first to learn to learn programs,
 an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:

@@

 \noindent \emph{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and go beyond it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} and Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}.

@@

 \section{Conclusion}
 We contributed a model, \system, that learns to program by
 bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the desired programs.  We believe that this
 integration of top-down symbolic representation and bottom-up neural
 networks -- both of them learned -- is a useful general approach for
 getting program induction off the ground.
